---
title: "February 2019: Visualizing my discogs collection"
author: ''
date: '2019-03-15'
slug: visualizing-discogs
---

In parts [one](/posts/discog-purrr/) and [two](/posts/more-discogging) of this series I did a whole lot of API pulling and data cleaning to get my [discogs](https://discogs.com/) collection into a tidy state `r emo::ji("pray")` Now I'm finally ready to do something with it!

I want to be able to explore my collection on a map (`r emo::ji("scream")`) and also see what styles of music I like, from where, and how that has changed over time.

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE, message = FALSE, warning = FALSE)
library(here)
library(dplyr)

collection_data <- readRDS(here("static", "data", "discogger", "collection_data.rds")) %>%
  as_tibble()
```

```{r}
collection_data
```

So, yes, I want to map my discogs collection all over the world `r emo::ji("globe")`

Pretty much everything I know about spatial data is from Jesse Sadler's amazing blog post, [Introduction to GIS with R](https://www.jessesadler.com/post/gis-with-r-intro/), so I'm pulling this code heavily from there.

First, so that we don't have legend fatigue, I'm going to lump the least common music styles together. My collection is fairly dominated by a few things:

```{r}
collection_data %>%
  count(style, sort = TRUE)
```

and while I'd love to specifically look at New Wave releases across the world, it just doesn't make sense for that grand total of 4.

```{r}
library(forcats)

collection_data <- collection_data %>%
  mutate(style = fct_lump(as_factor(style), 4))
```

Next, I'm converting my data frame into an `sf` object using the `long` and `lat` fields.

```{r}
library(sf)
library(dplyr)

points_sf <- collection_data %>%
  filter(!is.na(lat)) %>%
  st_as_sf(coords = c("long", "lat"), crs = 4326)

points_sf
```

In order to visualize those, I need a map of the world so I have something to plot on top of (I mean, I guess I don't *need* to use the actual earth as a reference point, but I think we'd all appreciate it if I did)

```{r}
library(rnaturalearth)

countries_sf <- ne_countries(scale = "medium", returnclass = "sf")
```

And then I can plot my collection! I'm using different colours for different music styles, and shapes for different formats.

To no surprise, the vast majority of my collection is from North America, with a real focus on the Pacific North West (I used to live in Vancouver `r emo::ji("umbrella")`) and Toronto/East Coast USA (there's just a lot of punk there, in general `r emo::ji("guitar")`).

```{r, out.width="100%"}
library(ggplot2)
library(paletteer)
library(plotly)

collection_plot <- ggplot() +
  geom_sf(data = countries_sf, fill = "white", size = 0.25, alpha = 0.5) +
  geom_sf(
    data = st_jitter(points_sf,
                     amount = 0.75),
    aes(color = style, shape = format,
        text = glue::glue('"{title}" by {artist_name}<br>{city}, {country}<br>{style} {format}')),
    alpha = 0.75,
    show.legend = FALSE,
    size = 2
  ) + 
  theme_bw() + 
  theme(legend.position = "none", 
        legend.title = element_blank(),
        axis.text = element_blank(), 
        axis.ticks = element_blank()) + 
  scale_color_paletteer_d("rcartocolor", "Pastel")

ggplotly(collection_plot, 
         tooltip = "text")
```

I'm also interested in the different ~eras~ of my music taste -- do I like different kinds of music from different times? You know how to add the time dimension to a plot?

Animation `r emo::ji("sunglasses")` `r emo::ji("shooting_star")`

Similar to spatial data, everything I know about animation is from one source: Thomas Lin Pedersen's [talk](https://resources.rstudio.com/rstudio-conf-2019/gganimate-live-cookbook) about the [`gganimate` package](https://github.com/thomasp85/gganimate) from RStudio conf.

I'm going to focus on North America, since that's where most of my information is from. In a maybe blasphemous move, I'm overlaying the American states and Canadian provinces and territories *over* the map of the world's countries `r emo::ji("grimace")`

```{r}
states_sf <- ne_states(country = c("Canada", "United States of America"), returnclass = "sf")

north_america_collection_plot <- ggplot() +
  geom_sf(data = countries_sf, fill = "white", size = 0.25) +
  geom_sf(data = states_sf, fill = NA, size = 0.25) + 
  geom_sf(
    data = st_jitter(points_sf %>% filter(year > 0),
                     amount = 0.75),
    aes(color = style, shape = format),
    alpha = 0.75,
    show.legend = "point",
    size = 3
  ) +
  theme_bw() + 
  theme(legend.title = element_blank(),
        legend.position = "bottom") + 
  guides(colour = guide_legend(override.aes = list(size=5, alpha = 1)),
         shape = guide_legend(override.aes = list(size=5, alpha = 1))) + 
  scale_color_paletteer_d("rcartocolor", "Pastel") + 
  coord_sf(xlim = c(-130, -65), ylim = c(23, 55), datum = NA)
```

Without animation, it's not bad.

```{r}
north_america_collection_plot
```

With animation it's *way* cooler.

```{r, cache=TRUE, out.width="100%"}
library(gganimate)

north_america_collection_plot + 
  transition_states(as.factor(year),
                    state_length = 3) + 
  ggtitle("{closest_state}") + 
  shadow_mark() 
```

I lived in the PNW from 2013 to 2017, and you can see a huuuge increase in music from there during that time. Pretty cool!

I think that's all I have `r emo::ji("information_desk_person")` Bye!

