---
title: here's what i know about tidyeval
author: ''
date: '2019-02-10'
slug: tidyeval
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE, message = FALSE)
```

there's no shortage of resources about tidyeval (i've listed some at the bottom), but this is a collection of what i know.

there is really no "why" here, or not much. i'm more of a "how" person, so ymmv on the usefulness.

i won't use mtcars or iris because i'm bored to death of them. let's use a dataset of toronto subway delays from 2018 ([available from toronto open data](https://www.toronto.ca/city-government/data-research-maps/open-data/open-data-catalogue/transportation/#917dd033-1fe5-4ba8-04ca-f683eec89761))

```{r read-data}
library(dplyr)

delays <- fs::dir_ls(here::here("static", "data", "ttc-delays", "delays")) %>%
  purrr::map_dfr(readxl::read_excel) %>%
  janitor::clean_names()

head(delays)
```

side note, but i can't believe it's that easy to read in 12 files and combine them. truly wild.

tidyeval time.

# one variable

let's say i want a function that returns the mean delay (`min_delay` is the delay, in minutes) based on a specific grouping, e.g. by `station`, maybe by `day`.

when writing the function, use `enquo()` to quote the variable, then `!!` to unquote it.

```{r grouped-mean-delay}
grouped_mean_delay <- function(df, group_var){
  group_var <- enquo(group_var)

  df %>%
    group_by(!!group_var) %>%
    summarise(mean_delay = mean(min_delay))
}
```

when i use the function, i can just call `grouped_mean_delay()` and pass it whatever variable i want to group by, without parentheses.

```{r use-grouped-mean}
delays %>%
  grouped_mean_delay(group_var = day)
```

# two variables, for two purposes

that's nice, but i probably don't always want the mean delay. what if i want the mean *gap* that the delay causes? the variable `min_gap` shows this -- e.g. if `min_gap` is 12, then that delay caused a 12 minute gap between trains at that station.

i don't really want to write a new function for every variable i might want to get the mean for, so it'd be nice to generalize `grouped_mean_delay()` to be a more general grouped mean.

you can do this the exact same way, and just add another argument for the variable you want the mean for.

```{r grouped-mean}
grouped_mean <- function(df, group_var, mean_var){
  group_var <- enquo(group_var)
  mean_var <- enquo(mean_var)
  
  df %>%
    group_by(!!group_var) %>%
    summarise(mean = mean(!!mean_var))
}

delays %>%
  grouped_mean(group_var = day, 
               mean_var = min_gap)
```

yes there are way to change the name of the output variable (i.e. so it's not just `mean`). [programming with dplyr](https://dplyr.tidyverse.org/articles/programming.html) talks about this, but i never really do it, so `r emo::ji("information_desk_person")`

# many variables, for the same purpose?

if i'm a curious person (i am), i probably want to be able to group by more than one thing at a time, e.g. by day *and* by subway line (`line`). 

there's a few ways you can do this.

## pass the dots

the first, which literally blew my mind the first time i saw it, uses `...`, and you pass the dots straight in when writing your function.

```{r}
grouped_mean_delay_2 <- function(df, ...){
  df %>%
    group_by(...) %>%
    summarise(mean_delay = mean(min_delay))
}

delays %>%
  grouped_mean_delay_2(day, line)
```

of course we have the added pleasure of the fact that this dataset isn't coded consistently (three variants of YU/BD!), but that's a topic for another post.

## pass the vars()

the thing about passing the dots is that those `...` are so mysterious. i definitely don't always write documentation for my functions, so it's nice to rely on named arguments that describe (even just a little!) what you should be throwing in there.

and sometimes it just doesn't work! in my mind, there are two kinds of verbs in dplyr:

1. verbs that take `...`, like `group_by()` and `select()`
2. (scoped) verbs that take `vars()`, like `mutate_at()` and `summarise_at()`

and you have to write your function using `vars()` a little differently, depending.

### verbs that take `...`

for verbs that take `...`, you got to just pass the dots. but you cannot just pass the `vars()`! if you want to use a named argument, and `vars()`, then you have to expand the variables back out using `!!!` (that's three bangs).

```{r}
grouped_mean_delay_3 <- function(df, group_vars){
  df %>%
    group_by(!!!group_vars) %>%
    summarise(mean_delay = mean(min_delay))
}

delays %>%
  grouped_mean_delay_3(group_vars = vars(day, line))
```

beauty.

### (scoped) verbs that take `vars()`

the `_at` scoped verbs, like `summarise_at()`, don't take `...` as an argument.

the `vars` argument of `summarise_at()` specifically says it "takes a list of columns generated by `vars()`" (and some other things). 

say we want the mean delay *and* the mean gap.

you can't pass the dots here.

```{r}
variable_mean_broken <- function(df, ...){
  df %>%
    summarise_at(..., mean)
}

delays %>%
  variable_mean_broken(min_delay, min_gap)
```

but you *can* just pass the `vars()`.

we need to pass in something that `summarise_at()` expects, specifically something that looks more like a `vars()` call. because `summarise_at()` *expects* something using `vars()`, we don't need to do anything to expand the variables out.

just like above, how because `group_by()` expects `...` arguments, we don't need to do anything to those dots. 

```{r}
variable_mean <- function(df, mean_vars){
  df %>%
    summarise_at(mean_vars, mean)
}

delays %>%
  variable_mean(mean_vars = vars(min_delay, min_gap))
```

beauty beauty.

# many variables, for many purposes?

i think this whole `vars()` thing really shines when you have many variables for many purposes. i knew about passing the dots, but i was like... how do you pass the dots... twice? jenny bryan's like yeah, you don't. 

you use `vars()`!

if i want to group by many variables *and* get the mean for many variables, then i can just throw a bunch of `vars()` in:

```{r}
general_grouped_mean <- function(df, group_vars, mean_vars){
  df %>%
    group_by(!!!group_vars) %>%
    summarise_at(mean_vars, mean)
}

delays %>%
  general_grouped_mean(group_vars = vars(line, day), 
                       mean_vars = vars(min_delay, min_gap))
```

the only thing here is that you *have* to use `vars()`, even if you're just passing one variable. like, this doesn't work:

```{r}
delays %>%
  general_grouped_mean(group_vars = day,
                       mean_vars = vars(min_delay, min_gap))
```

because the tidyeval method for passing in a single variable to `group_by()` is `enquo()` and then `!!` as described in the first bit,

and neither does this:

```{r}
delays %>%
  general_grouped_mean(group_vars = vars(day, line),
                       mean_vars = min_delay)
```

because `summarise_at()` requires a `vars()` call (or one of the other options described in the `.vars` argument).

this is still something i'm working out myself. how do i account for the possibility of a `vars()`, or just a variable on its own? is this what methods are for? s3? i totally know about that. for now, i will be a heavy user of the `vars()`, even when it's overkill, a la

```{r}
delays %>%
  general_grouped_mean(group_vars = vars(day),
                       mean_vars = vars(min_delay))
```

# passing functions

the last thing i know is something i *just* learned, because someone else asked about it on twitter. thanks [benjamin gowan](https://twitter.com/Benjaming_G/status/1094665581750509568)!

say i actually looked at my data and discovered the mean isn't a great measure for delays, so i want the median, too.

the way to do this is outside a function is:

```{r}
delays %>% 
  summarise_at(vars(min_delay, min_gap), 
               funs(mean, median))
```

so if you're writing a function, it looks like this:

```{r}
summary_by_var <- function(df, summary_vars, summary_funs){
  df %>%
    summarise_at(summary_vars, summary_funs)
}
```

if you're just doing one summary function, it's pretty easy to just pass right in

```{r}
delays %>%
  summary_by_var(vars(min_delay, min_gap),
                 median)
```

for multiple, you pass in a *list* of functions generated by `funs()`. i suggest you name them, otherwise your output will be ugly.

```{r}
delays %>%
  summary_by_var(vars(min_delay, min_gap),
                 funs(mean = mean, median = median))
```

you could actually use `list(mean = mean, median = median)` too, but i think the `funs()` case is clearer, since that's what `summarise_at()`'s `.funs` argument says it takes.

# putting it all together

now we can put everything together: group by multiple things and summarise multiple variables using multiple functions.

```{r}
grouped_summary <- function(df, group_vars, summary_vars, summary_funs){
  df %>%
    group_by(!!!group_vars) %>%
    summarise_at(summary_vars, summary_funs)
}

delays %>%
  grouped_summary(group_vars = vars(day, line),
                  summary_vars = vars(min_delay, min_gap),
                  summary_funs = funs(mean = mean,
                                      median = median))
```

wowwww.

# bye

there is a lot i don't know still, and i am happy to be pointed in the direction of other information, or told if i'm spreading misinformation! i especially would like to know how to be able to pass `vars()` OR just a single variable without `vars()`. do i write an `if` statement and then use `enquo()` and `!!`? i don't know! do you?

thanks to everyone that talks to me on twitter about tidyeval. there's a lot out there `r emo::ji("world")`

here is a collection of tidyeval resources that probably explain why things work the way they do:

* [programming with dplyr vignette by ?? the tidyverse team](https://dplyr.tidyverse.org/articles/programming.html) 
* [tidyeval book by lionel henry and hadley wickham](https://tidyeval.tidyverse.org/) (looks very much in dev, though i am so excited)
* [the 5 big ideas of tidyeval; short video by hadley wickham](https://www.youtube.com/watch?v=nERXS3ssntw)
* [lazy evaluation; rstudio conf talk by jenny bryan](https://resources.rstudio.com/rstudio-conf-2019/lazy-evaluation)